The Numerical Accuracy of Computing Electrostatic Interaction in the Inhomogeneous and Correlated Molecular Systems: for the Ewald Summation, SPME and Staggered Mesh Ewald Methods.

### 2012-07-13 longRange.hetero/alltext/publish/resubmit/lett.tex
fix bugs according to jctc template.

We have modified the introduction according to this point.
Please see the 4-th line of the second paragraph of the introduction:
\textbf{
  This problem can be solved by repeating and comparing
  the force calculation by
  using different parameter sets
  on a representative snapshot of the system,
  see eg. Ref.[16].
An other approach is the 
  in-depth theoretical study
of the error introduced by the algorithms, namely the
error estimate,
by which the error is described by a function of the working
parameters...
}\\

@@

We have added the comment (see the 12-th line of the 1st paragraph
of the introduction:
\textbf{It has been demonstrated that the SPME is actually
  a special case of the P3M method [6,7].}
The sentence ``We prove that the error is decomposed...'' has been
changed to \textbf{We show that the error may be approximately
  decomposed...}
\\

@@

We add new Fig. 9-12 to show how the error estimates
with bonded nearest neighbor approximations are able to catch the
trends of the real errors with various Ewald coefficient $\beta$, direct space
cut-off radius $r_c$, reciprocal space grid spacing $h$ and B-spline
interpolation order $n$.
% These figures show the error estimate
% with the bonded nearest neighbor approximation correctly catches the
% trend of changing working parameters.

In the revised manuscript,
we consider two types of the correlations:
(1) the bonded correlation
that originates from the rigid H-O
bond and H-O-H angle within one molecule,
and (2) the
non-bonded correlation or the
intermolecular correlation that originates from
a complex coupling of various intermolecular interaction.
Estimates for the bonded correlation
have already been developed by the old manuscript,
while the non-bonded estimate
is newly developed according to the reviewers' suggestion.
Please note that
both the bonded and non-bonded nearest neighbor approximations
require some uniformity of the correlation in the system:
the bonded estimate requires the uniformity of the water
direction preference, while the non-bonded estimate requires the uniformity
of the density distribution and the isotropicity of the system.
Otherwise the cost of the error estimate would be $\mathcal O(N^2\log N)$.
For the non-bonded correlation,  the nearest neighbor approximation 
is expressed in terms of the radial distribution
functions (RDFs). In example 3, a global RDF is not well defined,
so we calculated the RDF and nearest neighbor approximation
only for the liquid phase, see the dotted line in Fig.7. By adding
the non-bonded correlation, the quality of the error estimate
is further improved.
Also the convergence with respect to
the neighboring range $R_g$ is investigated (Fig.8). 
\\

@@

In the revised manuscript, the correlation error are considered
in two types: 
(1) the bonded correlation
that originates from the rigid H-O
bond and H-O-H angle within one molecule,
and (2) the
non-bonded correlation or the
intermolecular correlation that originates from
a complex coupling of various intermolecular interaction.
Estimates for the bonded correlation
have already been developed by the old manuscript,
while the non-bonded estimate
is newly developed according to the reviewers' suggestion.
Please note that
both the bonded and non-bonded nearest neighbor approximations
require some uniformity of the correlation in the system:
the bonded estimate requires the uniformity of the water
direction preference, while the non-bonded estimate requires the uniformity
of the density distribution and the isotropicity of the system, which
are also the prerequisites for reducing $\rho^{(2)}$ to $g(r)$.
%Otherwise the cost of the error estimate would be $\mathcal O(N^2\log N)$.
For the non-bonded correlation,  the nearest neighbor approximation 
is expressed in terms of the radial distribution
functions (RDFs) in the new Eqn. (21) and (22).
In example 3, a global RDF is not well defined
(because the system is inhomogeneous),
so we calculated the RDF and nearest neighbor approximation
only for the liquid phase, see the dotted line in Fig.7. By adding the
non-bonded correlation, the quality of the error estimate
is further improved.
Also the convergence with respect to
the neighboring range $R_g$ (the $r_c$ in the reviewer's comment)
is investigated (Fig.8). 
\\

@@

As far as we concerned,
direct error estimates with the higher order terms are not easy.
Therefore, we investigate this issue in another way:
we compare the estimate (with non-bonded nearest neighbor
approximation) with the real error of the randomized COM test
(see the description of this test in page 25).
Since the only approximation made in the estimate is the leading order
approximation, the difference between the estimate and the real error
of randomized COM test is actually the contribution from the higher
order terms discarded by (48) and (53).
In Fig. 9-12,
we check this difference in a wide parameter range, and see in most
cases the estimate overlaps with the real error, which means
the leading order is a good approximation. 
However,
we do observe exceptions
at large $\beta$ when the grid spacing is $h=0.248\:\textsf{nm}$,
see Fig.11, and
at small $\beta$ when the interpolation order is $n=4$,
see Fig.12.
That means the higher order terms may play a role in the staggered
mesh interpolation errors. It also explains the unexpected discrepancy
between the staggered mesh ik- and analytical differentiation
at small $\beta$, $n=4$.
\\

@@

We define the force error by inserting a testing charge into the
considered system, and comparing the calculated and exact force on
it. In the considered water system (Example 3), the subtle difference
between this setting to the realistic case is that the testing charge
has no correlation with the charges in the system, while any real
charge is bonded to two neighboring charges (via H-O bond and H-O-H
angle), as is illustrated by the new Fig.1.
% Therefore, the error estimate does not consider the
% correlation between a charge and the its bonded neighbors, but it does
% consider all bonding correlations within any neighboring molecule (see
% new Fig.1 of the paper).
This definition is actually consistent with the usually
setting of bonded pair atoms, as mentioned by the reviewer, the
electrostatic interaction between them is not calculated.
Throughout this paper, we remove this  interaction from the
standard long-range algorithms.

@@

As far as we concerned, it is not possible to test this point
directly, but we think it is a likely explanation to the observed
phenomenon.  Consider the extreme case: within one molecule, the
hydrogen and oxygen atoms locate at the same point (strongest
correlation), then the positive and negative charges cancel, so the
error vanishes. In the case of TIP3P water model, the H-O distance is 0.1~nm,
which is much smaller than the typical
center-of-mass distance between two neighboring molecules (0.31~nm),
therefore, the bonded correlation tends to neutralize the molecule
and reduce the force error.
We have added this point to the paper.
\\

>>>>

We have modified the introduction according to this point.
Please see the 4-th line of the second paragraph of the introduction:
\textbf{
  This problem can be solved by  comparing
  the forces calculated by
  using different parameter sets
  on a representative snapshot of the system,
  see e.g. Ref.[16].
An other approach is the 
  in-depth theoretical study
of the error introduced by the algorithms, namely the
error estimate,
by which the error is described by a function of the working
parameters...
}\\

@@

We have added the comment (see the 12-th line of the 1st paragraph
of the introduction:
\textbf{It has been demonstrated that the SPME is actually
  a special case of the P3M method [6,7].}
The sentence ``We prove that the error is decomposed...'' has been
changed to \textbf{We show that the error may be approximately
  decomposed...}, please see the 4-th line of the third paragraph of the
introduction.
\\

@@

We add new Fig. 9-12 to show how the error estimates
with bonded nearest neighbor approximations are able to catch the
trends of the real errors with various Ewald coefficient $\beta$, direct space
cut-off radius $r_c$, reciprocal space grid spacing $h$ and B-spline
interpolation order $n$.
Please also see the discussions added
from the last line of page 29.
% These figures show the error estimate
% with the bonded nearest neighbor approximation correctly catches the
% trend of changing working parameters.

In the revised manuscript,
we consider both of the two types of the correlations:
(1) the bonded correlation
that originates from the rigid H-O
bond and H-O-H angle within one molecule,
and (2) the
non-bonded correlation or the
intermolecular correlation that originates from
a complex coupling of various intermolecular interaction.
Estimates for the bonded correlation
have already been developed by the old manuscript,
while the non-bonded estimate
is newly added according to the reviewers' suggestion.
Please note that
both the bonded and non-bonded nearest neighbor approximations
require some uniformity of the correlation in the system:
the bonded estimate requires the uniformity of the water
direction preference (as has been discussed in the old manuscript),
while the non-bonded estimate requires the uniformity
of the density distribution and the isotropicity of the system.
Otherwise the cost of the error estimate would be $\mathcal O(N^2\log N)$.
For the non-bonded correlation,  the nearest neighbor approximation 
is expressed in terms of the radial distribution
functions (RDFs), see the new Eq. (21) and (22). In example 3, a global RDF is not well defined,
so we calculated the RDF and nearest neighbor approximation
only for the liquid phase, see the dotted line in Fig.7. By adding
the non-bonded correlation, the quality of the error estimate
is further improved.
Also the convergence with respect to
the neighboring range $R_g$ is investigated, see Fig.8.
Relavant changes to the manuscript
have been made on page 9,
11 (4-th line from the bottom), 12, 29 (12-th line). All changes are marked 
in red for convenience.
\\

@@

In the revised manuscript, the correlation error are considered
in two types: 
(1) the bonded correlation
that originates from the rigid H-O
bond and H-O-H angle within one molecule,
and (2) the
non-bonded correlation or the
intermolecular correlation that originates from
a complex coupling of various intermolecular interaction.
Estimates for the bonded correlation
have already been developed by the old manuscript,
while the non-bonded estimate
is newly developed according to the reviewers' suggestion.
Please note that
both the bonded and non-bonded nearest neighbor approximations
require some uniformity of the correlation in the system:
the bonded estimate requires the uniformity of the water
direction preference (as has been discussed in the old manuscript),
while the non-bonded estimate requires the uniformity
of the density distribution and the isotropicity of the system, which
are also the prerequisites for reducing $\rho^{(2)}$ to $g(r)$.
Otherwise the cost of the error estimate would be $\mathcal O(N^2\log N)$.
For the non-bonded correlation,  the nearest neighbor approximation 
is expressed in terms of the radial distribution
functions (RDFs) in the new Eqn. (21) and (22).
In example 3, a global RDF is not well defined
(because the system is inhomogeneous),
so we calculated the RDF and nearest neighbor approximation
only for the liquid phase, see the dotted line in Fig.7. By adding the
non-bonded correlation, the quality of the error estimate
is further improved.
Also the convergence with respect to
the neighboring range $R_g$ (the $r_c$ in the reviewer's comment)
is investigated (Fig.8). 
Relavant changes to the manuscript
have been made on page 9,
11 (4-th line from the bottom), 12, 29 (12-th line). All changes are marked 
in red for convenience.
\\

@@

As far as we concerned,
direct error estimates with the higher order terms are not easy.
Therefore, we investigate this issue in another way:
we compare the estimate (with bonded nearest neighbor
approximation) with the real error of the randomized COM test
(see the description of this test at the 4th line from bottom of page 28).
Since the only approximation made in the estimate is the leading order
approximation, the difference between the estimate and the real error
of randomized COM test is actually the contribution from the higher
order terms discarded by (48) and (53).
In Fig. 9-12,
we check this difference in a wide parameter range, and see in most
cases the estimate overlaps with the real error, which means
the leading order is a good approximation. 
However,
we do observe exceptions
at large $\beta$ when the grid spacing is $h=0.248\:\textsf{nm}$,
see Fig.11, and
at small $\beta$ when the interpolation order is $n=4$,
see Fig.12.
That means the higher order terms may play a role in the staggered
mesh interpolation errors. It also explains the unexpected discrepancy
between the staggered mesh ik- and analytical differentiation
at small $\beta$, $n=4$.
Relavant changes to the manuscript
have been made on page 28 (5-th line from the bottom), 32 (5-th line from the bottom).
\\

@@

We define the force error by inserting a testing charge into the
considered system, and comparing the calculated and exact force on
it. In the considered water system (Example 3), the subtle difference
between this setting to the realistic case is that the testing charge
has no correlation with the charges in the system, while any real
charge is bonded to two neighboring charges (via H-O bond and H-O-H
angle), as is illustrated by the new Fig.1.
% Therefore, the error estimate does not consider the
% correlation between a charge and the its bonded neighbors, but it does
% consider all bonding correlations within any neighboring molecule (see
% new Fig.1 of the paper).
This definition is actually consistent with the usually
setting of bonded pair atoms, as mentioned by the reviewer, the
electrostatic interaction between them is not calculated.
Throughout this paper, we remove this  interaction from the
standard long-range algorithms.  This discuss has been added to
the manuscript, see from the 3rd line under Eq.20.

@@

As far as we concerned, it is not possible to test this point
directly, but we think it is a likely explanation to the observed
phenomenon.  Consider the extreme case: within one molecule, the
hydrogen and oxygen atoms locate at the same point (strongest
correlation), then the positive and negative charges cancel, so the
error vanishes. In the case of TIP3P water model, the H-O distance is 0.1~nm,
which is much smaller than the typical
center-of-mass distance between two neighboring molecules (0.31~nm),
therefore, the bonded correlation tends to neutralize the molecule
and reduce the force error.
We have added this point to the paper, see line 10 on page 28.
\\
